---
title: "Arrival_Cleaning_Part2"
author: "Carole Shaw"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

## Purpose

This script is designed to put arrival data together and to distinguish data collected for treatment prior to FHCC arrival, arrivals associated with FHCC care, patients who have not had any arrival data entered yet, and arrivals that not associated with a currative treatment. The plan is to add resulting data from this routine to SQL Server updating it with the weekly workflow so that the data will be consistently compiled and updated over time.

The data created here are joined with other source data, such as transplant data, epic report data, and eventually with CORTEX downloads during the creation of the frozen data.

## TREATMENT VARIATION -- FUTURE TREATMENT STANDARDIZATION

```{r}
allarrival_ <- get(backup_name.2)
FHCC_Care <- get(backup_name.2) |>
  filter(category=='FH Treatment' | category=='No Treatment') |>
  filter(!is.na(treatment)) |>
  group_by(category, treatment, treatmentintensity_calc, treatmentbackbone_calc, treatmentbcrabl_tki, treatmentflt3_tki) |>
  summarize(instances = n(),
            .groups='drop') |>
  arrange(desc(instances))


# Take a look at the unique treatment texts by category "FH Treatment"
Currative_Treatment_at_FHCC <- get(backup_name.2) |>
  filter(category=='FH Treatment' & currativerx_fhcc & !is.na(treatment) & str_detect(treatment, '(?i)died', negate=TRUE)) |>
  group_by(category, treatment, treatmentintensity_calc, treatmentbackbone_calc, treatmentbcrabl_tki, treatmentflt3_tki) |>
  summarize(instances = n(),
            .groups='drop') |>
  arrange(desc(instances))

identical(allarrival$treatment_intensity, allarrival_$treatment_intensity)

# Take a look at the unique treatment texts by category "Prior Treatment Outside"
Currative_Treatment_Outside <- get(backup_name.2) |>
  filter(category=='Prior Treatment Outside' & currativerx_any & !is.na(treatment)) |>
  group_by(category, treatment, treatmentintensity_calc, treatmentbackbone_calc, treatmentbcrabl_tki, treatmentflt3_tki) |>
  summarize(instances = n(),
            .groups='drop') |>
  arrange(desc(instances))

make_sheet(FHCC_Care, scrapdir, sheet='data', 'FHCC_Care', time=TRUE, overwrite = TRUE)


```

## RG Number Identification

Maybe this should be it's own project that deals more with lining up protocol names and patient enrollment in protocols and is based on CTMS downloads as well as frozen.glacier data.

```{r}
protocol_lkup <- sql_get_table('frozen.protocol_lkup') |>
  filter(`Protocol.No.`!='RG1001098')

xx <- inner_join(
  get(backup_name.2) |>
    select(any_of(idlist), row_number, treatment, arrivaltable, protocol, starts_with('rg')),
  protocol_lkup |> 
    mutate(rgnumber_ = `Protocol.No.`,
           treatment = IRB), 
    select(rgnumber_, `Protocol.No.`, `IRB`, Title, Short.Title, `PI.Name`),
  by = 'treatment') |>
  select(any_of(idlist), row_number, treatment, arrivaltable, starts_with('rg'),
         rgnumber, protocol, `Protocol.No.`, IRB, Title, Short.Title, `PI.Name`)

yy <- inner_join(
  get(backup_name.2) |>
    select(any_of(idlist), row_number, treatment, arrivaltable, protocol, starts_with('rg')),
  protocol_lkup |> 
    mutate(rgnumber_ = `Protocol.No.`,
           protocol = IRB), 
    select(rgnumber_, `Protocol.No.`, IRB, Title, Short.Title, `PI.Name`),
  by = 'protocol') |>
  select(any_of(idlist), row_number, treatment, arrivaltable, starts_with('rg'),
         rgnumber, protocol, `Protocol.No.`, IRB, Title, Short.Title, `PI.Name`)


zz <- get(backup_name.2) |>
  select(any_of(idlist), row_number, treatment, arrivaltable, starts_with('rg')) |>
  inner_join(
    protocol_lkup |>
      mutate(treatment = `Short.Title`,
             rgnumber_ = `Protocol.No.` ) |>
      select(IRB, treatment, rgnumber_, `Protocol.No.`, Title, `Short.Title`, `PI.Name`),
    by = "treatment"
  ) |>
  group_by(across(c(any_of(idlist), row_number, treatment, arrivaltable, rgnumber_, `Protocol.No.`, Title, `Short.Title`, `PI.Name`))) |>
  summarise(
    IRB_aliases = str_c(unique(IRB), collapse = "; "),
    .groups = "drop"
  )

RG_Number_Updates <- bind_rows(xx, yy, zz) |> 
  filter(is.na(rgnumber)) |> 
  mutate(rgnumber = rgnumber_) |>
  select(any_of(idlist), row_number, treatment, arrivaltable, 
         rgnumber, `Protocol.No.`, Title, Short.Title, `PI.Name`) |>
  distinct() |>
  build_arrivalupdates()

make_sheet(RG_Number_Updates, rcimportdir, filename='RG_Number_Updates', null_as_blank = TRUE)
```


## FHCC FIRST/LAST ARRIVAL

Maybe we should first eliminate any arrivals that are dups?

```{r}
# identify earliest FHCC arrival
neatfreq(allarrival$category)

allarrival <- get(backup_name.2)

first_fhcc_arrival <- filter(allarrival, str_detect(category, '(?i)outside', negate=TRUE)) |>
  mutate(min_date = pmin(arrivaldate, treatmentdate, na.rm = TRUE)) |>
  select(recordid, min_date) |>
  filter(!is.na(min_date)) |>
  group_by(recordid) |>
  summarise(firstfhccarrivaldate = min(min_date)  # Get min_date for each group
    , .groups = 'drop')

last_fhcc_arrival <- filter(allarrival, str_detect(category, '(?i)outside', negate=TRUE)) |>
  mutate(max_date = pmax(arrivaldate, treatmentdate, na.rm = TRUE)) |>
  select(recordid, max_date) |>
  filter(!is.na(max_date)) |>
  group_by(recordid) |>
  summarise(lastfhccarrivaldate = max(max_date)         # Get max_date for each group            
    , .groups = 'drop')


first_known_arrival <- allarrival |>
  mutate(min_date = pmin(arrivaldate, treatmentdate, na.rm = TRUE)) |>
  select(recordid, min_date) |>
  filter(!is.na(min_date)) |>
  group_by(recordid) |>
  summarise(firstknownarrivaldate = min(min_date)  # Get min_date for each group
    , .groups = 'drop')

last_known_arrival <- allarrival |>
  mutate(max_date = pmax(arrivaldate, treatmentdate, na.rm = TRUE)) |>
  select(recordid, max_date) |>
  filter(!is.na(max_date)) |>
  group_by(recordid) |>
  summarise(lastknownarrivaldate = max(max_date)         # Get max_date for each group            
    , .groups = 'drop')

allarrival <- allarrival |>
  left_join(first_fhcc_arrival) |>
  left_join( last_fhcc_arrival) |>
  left_join(first_known_arrival) |>
  left_join( last_known_arrival)

backup_name.3 <- paste0('allarrival_', as.character(now()))
assign(backup_name.3, allarrival)
allarrival <- get(backup_name.3)


cat('Done\n')

```

## INDUCTION ARRIVALS

```{r}
allarrival <- get(backup_name.3) |> 
  arrange(recordid, arrivaldate) |>
  mutate(allarrivalrow = row_number()+10000) |> 
  select(allarrivalrow, any_of(idlist), everything())

# allarrival <- sql_get_table('frozen.allarrival')

aml_reasons <- c(">20%", ">10%", "t(15;17) (APL)", "t(8;21)", "inv(16)",
                 "NPM1+", "EMD", "AUL", "MPAL", "tp53", "BPDCN", "Other WHO AML")

arrival_induction_scores <- allarrival |>
  select(allarrivalrow, any_of(idlist), 
         currativerx_fhcc, currativerx_any, 
         category, arrivalnum, arrivaltable,
         dxdate, arrivaldate, arrivaltype, arrivalchar, rxline, arrivalreason, 
         arrivalmorphdate, arrivalmorph, arrivalflowdate, arrivalflow,
         a_treatmentlocation, treatmentdate, treatment, 
         starts_with(c('treatment', 'first', 'last')), special_population, everything()) |>
  mutate(
    num_arrivalmorphnum = as.numeric(str_extract(arrivalmorph, "\\d+(?:\\.\\d+)?")),
    num_arrivalflownum  = as.numeric(str_extract(arrivalflow,  "\\d+(?:\\.\\d+)?")),
    days_dx_to_arrival  = as.integer(arrivaldate - dxdate),
    days_arrival_to_rx  = as.integer(treatmentdate - arrivaldate),
    days_dx_to_rx       = as.integer(treatmentdate - dxdate),
    score_1             = as.integer(str_detect(arrivaltype, 'New')) * 2,
    score_2             = as.integer(rxline=='Ind') * 2,
    score_3             = as.integer(currativerx_any),
    score_4             = as.integer(str_detect(arrivaltype, 'New')) + as.integer(rxline=='Ind'),
    score_5             = as.integer(currativerx_any) + as.integer(str_detect(arrivaltype, 'New')) + as.integer(rxline=='Ind'),
    score_6             = as.integer(num_arrivalmorphnum >= 10 | num_arrivalflownum >= 20),
    score_7             = as.integer(arrivalreason %in% aml_reasons),
    score_8             = as.integer(dxdate <= arrivaldate),
    score_9             = as.integer(arrivaldate <= treatmentdate),
    score_10            = as.integer(days_dx_to_arrival <= 30),
    score_11            = as.integer(days_dx_to_rx <= 30),
    score_12            = as.integer(arrivaltable=="induction") * 3,
    score_13            = as.integer(currativerx_any  & is.na(currativerx_fhcc) & str_detect(firstinduction, "outside")) * 2,
    # (Optional fix) if you meant “has a treatment date”:
    # score_14         = as.integer(currativerx_any & is.na(currativerx_fhcc) & !is.na(treatmentdate)) * 2,
    score_14            = as.integer(currativerx_any  & is.na(currativerx_fhcc) & 'treatmentdate'>0) * 2,
    score_anti_1        = as.integer(str_detect(arrivaltype, '(?i)salvage|mrd')) * -1,
    score_anti_2        = as.integer(str_detect(rxline, '(?i)s\\d+')) * -1,
    score_zero_not_curr = as.integer(currativerx_fhcc | is.na(currativerx_fhcc)),
    score_zero_pre_dx   = as.integer(((days_dx_to_arrival < 90) | is.na(days_dx_to_arrival)) &
                                     ((days_dx_to_rx      < 90) | is.na(days_dx_to_rx)) &
                                     !(is.na(days_dx_to_arrival) & is.na(days_dx_to_rx))),
    score_sub           = rowSums(pick(starts_with("score_")), na.rm = TRUE),
    score_factored      = score_sub * score_zero_not_curr * score_zero_pre_dx,
    confidence_score    = score_factored
  ) |>
  arrange(recordid, ptmrn, desc(confidence_score), arrivaldate) |>
  group_by(ptmrn) |>
  mutate(
    arrival_cnt           = n(),
    currativerx_fhcc_cnt  = sum(coalesce(currativerx_fhcc, FALSE)),  # logical-safe count
    arrseq                = row_number(),

    # Band each row by confidence threshold
    induction_band = dplyr::case_when(
      confidence_score >= 14      ~ "strong confidence",
      confidence_score %in% 12:13 ~ "confident",
      confidence_score %in% 10:11 ~ "not confident",
      .default = NA_character_
    ),

    # Candidate rows are those above threshold
    .induction_candidate = !is.na(induction_band),

    # Keep ONLY the first candidate per patient as TRUE; everything else FALSE
    induction = .induction_candidate & cumsum(.induction_candidate) == 1,

    # Descriptive label only on the TRUE row
    induction_confidence = if_else(induction, induction_band, NA_character_)
  ) |>
  ungroup() |>
  select(
    allarrivalrow, any_of(idlist), arrseq,
    induction, induction_confidence, confidence_score,
    ends_with('_cnt'),
    starts_with(c('curr', 'num_', 'days', 'score_anti', 'score_zero')),
    everything()
  )

induction.pool <- arrival_induction_scores |>
  filter(str_detect(rxline, '(?i)(consult|prior)', negate=TRUE)) |>
  filter(arrivaltable!='patient_list') |>
  arrange(recordid, desc(confidence_score), arrivaldate) |>
  group_by(across(all_of(idlist))) |>
  slice(1) |>
  ungroup() |>
  mutate(ptlastname = toupper(ptlastname),
         inductionarrivallocation=a_treatmentlocation,
         inductionarrivaltreatmentdate=treatmentdate,
         inductionarrivaltreatment=treatment,
         inductionarrivalintensity=treatmentintensity_calc,
         inductionarrivalbackbone=treatmentbackbone_calc,
         inductionarrivalbcrabl=treatmentbcrabl_tki,
         inductionarrivalflt3=treatmentflt3_tki,
         inductionarrivaltreatmentmap=case_when(
           confidence_score==0  ~ 'Not induction',
           confidence_score>=14 ~ paste0('Induction (strong confidence=',confidence_score,')'),
           confidence_score>=12 ~ paste0('Induction (confident=',confidence_score,')'),
           confidence_score>=10 ~ paste0('Induction (not confident=',confidence_score,')'),
           confidence_score<10  ~ paste0('Undertermined (confidence=',confidence_score,')'),
           arrivaltable == 'patient_list' ~ 'Pre-Abstraction',
           TRUE~'Fallout'),
         inductionarrivaldxtorxdays=days_dx_to_rx,
         inductionarrivaldate=arrivaldate,
         inductionarrivalnum=arrivalnum,
         inductionarrivaltable=arrivaltable,
         inductionarrivaltype=arrivaltype, # should be ND, but isn't always
         inductionarrivalrxline=rxline, # should be Ind, but isn't always
         inductionarrivalmorphdate=arrivalmorphdate,
         inductionarrivalmorph=arrivalmorph,
         inductionarrivalflowdate=arrivalflowdate,
         inductionarrivalflow=arrivalflow
         ) |>
  select(allarrivalrow, any_of(idlist), dxdate, starts_with('inductionarrival'), everything())

rcexportdf <- induction.pool |>
  mutate(is_inductionarrival=currativerx_any & confidence_score >= 10
         , ptlastname = toupper(ptlastname)) |>
  select(is_inductionarrival, any_of(idlist), 
         starts_with(c("inductionarrival")),
         firstfhccarrivaldate, firstknownarrivaldate,
         lastfhccarrivaldate,  lastknownarrivaldate, 
         -inductionarrivaldxtorxdays, -inductionarrivaltreatmentmap) |>
  mutate(across(where(is.Date), ~ format(., "%Y-%m-%d"))) |>
  mutate(across(everything(), ~ ifelse(is.na(.), "", .)))

rcexportdf_keep <- rcexportdf |> 
  filter(is_inductionarrival) |>
  select(any_of(idlist), is_inductionarrival, starts_with('inductionarrival'))

allarrival <- allarrival |> left_join(rcexportdf_keep, by=idlist)

indarrival <- allarrival %>%
  group_by(recordid, ptmrn, ptlastname) %>%
  arrange(desc(is_inductionarrival)) %>%     # put TRUE first
  slice_head(n = 1) %>%                      # keep only the first row per group
  ungroup() %>%
  mutate(is_inductionarrival=as.integer(is_inductionarrival),
         ptlastname=toupper(ptlastname))  |>
  filter(inductionarrivaltable!='patient_list') |>
  select(any_of(idlist), inductionarrivalcategory=category, starts_with("inductionarrival"))

make_sheet(indarrival, rcimportdir, filename="induction_import", overwrite = TRUE, null_as_blank = TRUE)

backup_name.4 <- paste0('allarrival_', as.character(now()))
assign(backup_name.4, allarrival)
allarrival <- get(backup_name.4)

```

## PUTTING IT ALL TOGETHER WITH LAG/LEAD VARS

```{r}

allarrival <- get(backup_name.4)

allarrival <- allarrival |>
  arrange(recordid, arrivaldate, treatmentdate, arrivalnum, arrivaltable) |>
  select(-row_number) |>
  mutate(row_number = row_number()+10000) |>
  group_by(recordid, ptmrn, ptlastname) |>
  mutate(arrivalsequence = row_number(),
         arrivalcnt = n(),
         prevarrivaldate=lag(arrivaldate),
         prevarrivalcategory=lag(arrivalcategory),
         prevtreatmentdate=lag(treatmentdate),
         prevtreatment=lag(treatment),
         prevrxline=lag(rxline),
         nextarrivaldate=lead(arrivaldate),
         nextarrivalcategory=lead(arrivalcategory),
         nexttreatmentdate=lead(treatmentdate),
         nexttreatment=lead(treatment),
         nextrxline=lead(rxline)
         ) |>
  ungroup() |>
  select(row_number, arrivalsequence, any_of(idlist), primary_subject_id
         , category, arrivalcnt, firstfhccarrivaldate, lastfhccarrivaldate
         , starts_with(c('arrival', 'prev', 'next')), everything())
```

## QC

### arrival after treatment started

```{r}

arrivaldate_after_treatmentdate <- allarrival |>
  filter(str_detect(category, '(?i)(no\\streatment|without)', negate=TRUE)) |>
  mutate(days_arrival_to_rx = as.integer(treatmentdate-arrivaldate),
         days_morph_to_rx   = as.integer(treatmentdate-arrivalmorphdate)) |>
  filter(days_arrival_to_rx < -7 & days_morph_to_rx < -7) |>
  select(any_of(idlist), starts_with('days_'), category, 
         arrivaldate, rxline, arrivalmorphdate, 
         treatmentdate, treatment, 
         nextarrivaldate, nextrxline)

make_sheet(arrivaldate_after_treatmentdate, qcdir, 'data', 'suspicious arrival date')
```

## Report Missing Arrival Sequence number or Arrival Date

Very hard to get arrivals in order without this information, luckily not very many are missing these. When patients are entered into REDCap they are assigned a sequence number on the top of each arrival form. This is useful in the interface if we need to change the order in which arrivals occur. This becomes especially handy if a treatment attempt was missed during abstraction and needs to be inserted.\
\
The consequence is that sometimes we forget to adjust all of the sequence numbers associated with the various arrivals for the patient. When this is the case this code checks for any discrepancies and creates a report.\
\
It also checks for arrival date, missing arrival dates also make it difficult to tell in what order events took place.

```{r eval=FALSE, include=FALSE}
# identify errors
missing_arrival_date <- allarrival |> 
  filter(str_detect(arrivaltable, '^arrival') & 
           recordid > 1000 & 
           ptlastname != 'PATIENT' &
           str_detect(special_population,'(Older AML.*|Mucor.*)', negate=TRUE) &
           (is.na(arrivaldate) | arrivalnum == "")) |>
  select(  any_of(idlist) #, arrival_rectime
         , arrivaltable=orig_arrivaltable, arrivaldate, arrivaltype, rxline, arrivalnote
         , treatmentdate, treatmentnote, special_population )
  

if (nrow(missing_arrival_date) > 0){
  # Create excel for data cleaning
  make_sheet(missing_arrival_date, scrapdir, 'no_arrival_date', 'missing_arrival_date', 'xlsx', overwrite=TRUE )
  # Add to table of things to correct
  addrow('todo_df', scrapdir, 'missing_arrival_date.xlsx', 'Review for arrival date missing and arrival number order')
  cat('\nReport Missing Arrival Sequence number or Arrival Date')
} else {
  cat('\nNo Missing Arrival Sequence number or Arrival Date')
}
```

## SAVE TO SQLSERVER

Copy tables to SQL Server. The table **allarrival** contains one record for each arrival entered in arrival1 to 11, a record for any patient entered in the patient list who have not yet been abstracted and a record from the former diagnosis REDCap if the patient had outside induction.

There are several fields that are key to navigating sub-populations in this table:

-   category

Frequency Table for allarrival\$category

| Value                             | Frequency | Percentage |
|:----------------------------------|:---------:|:----------:|
| allarrival\$category              |           |     \%     |
| FH Treatment                      |   4589    |   71.65%   |
| Prior Treatment Outside           |    889    |   13.88%   |
| No Treatment                      |    748    |   11.68%   |
| Outside Arrival without Treatment |    179    |   2.79%    |

-   currativerx_fhcc (patients who got treatment at our center)

-   currativerx_any (patients who got treatment anywhere)

```{r to_sqlserver}
# Resize accommodate structure of all the records so that when we append columns won't be customized in width by the sql_insert_df()
tempsch <- 'scratch'
frosch  <- 'frozen'
sql_drop_tbl(tempsch, 'temp')

# customize the column width based on this concatenated df
sql_insert_df(allarrival, 'scratch.temp')

# make empty copies of the structure
sql_copy_tbl(frosch, 'allarrival', tempsch, 'temp', data=FALSE)

# insert allarrivals, allnotreat, and allprior twice, normal and again with a date suffix
sql_insert_df(allarrival, 'frozen.allarrival', date_suffix='%Y%m', overwrite = FALSE) 
# create or update table description
sql_desc_tbl(paste0(frosch, '.allarrival'), desc='All arrivals.  The columns category, currativerx_fhcc, and currativerx_any are helpful in determining sub populations')

# insert induction arrival data into frozen schema on SQL Server
sql_insert_df(indarrival, 'frozen.indarrival', date_suffix='%Y%m', overwrite = TRUE) 
# create or update table description
sql_desc_tbl(paste0(frosch, '.indarrival'), desc='Arrivals that are calculated to be first AML induction')


# print table descriptions after addition/update
sql_desc_tbl(paste0(frosch, '.allarrival'))

cat('SQL Server updated with allarrival, and dated copy')
```

## ---- OBSOLETE ----

## ARRIVAL SPECIFIC INTENSITY AND BACKBONE

I think this code has been improved in the chunk:  "CATEGORIZE INTENSITY, BACKBONE, TKI's"

```{r eval=FALSE, include=FALSE}
# pivoting wide to get this into a format for upload to REDCap
uploaddf <- allarrival |>
  filter(str_detect(arrivaltable, "arrival\\d+")) |>
  mutate(
    arrival_num = parse_number(arrivaltable),
    ptlastname  = toupper(ptlastname)
  ) |>
  filter(dplyr::between(arrival_num, 1, 11)) |>
  select(
    any_of(idlist),
    arrival_num,
    treatment,
    treatmentintensity = treatment_intensity,
    treatmentbackbone  = treatment_backbone
  ) |>
  # Normalize empties to NA so we can reason cleanly
  mutate(
    treatment           = dplyr::na_if(trimws(treatment), ""),
    treatmentintensity  = dplyr::na_if(trimws(treatmentintensity), ""),
    treatmentbackbone   = dplyr::na_if(trimws(treatmentbackbone), "")
  ) |>
  # Apply REDCap upload rules:
  # - If no treatment: mapped fields should be blank (leave NA; pivot will fill "")
  # - If treatment exists but mapping is missing: use "no map"
  mutate(
    treatmentintensity = dplyr::case_when(
      is.na(treatment)                      ~ NA_character_,
      !is.na(treatment) & is.na(treatmentintensity) ~ "no map",
      TRUE ~ treatmentintensity
    ),
    treatmentbackbone = dplyr::case_when(
      is.na(treatment)                      ~ NA_character_,
      !is.na(treatment) & is.na(treatmentbackbone)  ~ "no map",
      TRUE ~ treatmentbackbone
    )
  ) |>
  tidyr::pivot_wider(
    id_cols     = any_of(idlist),
    names_from  = arrival_num,
    values_from = c(treatmentintensity, treatmentbackbone),
    names_glue  = "{.value}_calc{arrival_num}",
    names_sort  = TRUE,
    # Prefer a real value over "no map"; fall back to "no map"; else blank via values_fill
    values_fn   = ~ {
      x <- .x[!is.na(.x) & .x != ""]
      if (length(x) == 0) return("")
      real <- x[x != "no map"]
      if (length(real) > 0) real[[1]] else x[[1]]
    },
    values_fill = ""
  )

make_sheet(uploaddf, rcimportdir, filename='upload_arrival_treatment_summary', overwrite = TRUE)
```

### ERROR OUTSIDE treatments dated *before* the first arrival recorded at FHCC

Answers the question: Are there arrivals dated after first FHCC arrival/treatment that we have indicated as "OUT" for location?

Then produces an excel document in the scrap directory with records that seem fishy.

```{r eval=FALSE, include=FALSE}
# Are there prior arrivals dated after first FHCC arrival/treatment

# Have staff review any of these where arrival date is after first FHCC arrival, but Rx date is not
late_outside_arrival <- allprior |>
  mutate(  priorarrivaldate    = arrivaldate
         , priortreatmentdate  = treatmentdate
         , afterfirstarrival   = priorarrivaldate   > firstfhccarrivaldate
         , afterfirsttreatment = priortreatmentdate > firstfhccarrivaldate
         , reviewreason = paste('Arrival date',arrivaldate
                                ,'is after first arrival on',firstfhccarrivaldate)) |>  
  filter(  afterfirstarrival & !afterfirsttreatment) |>
  select(  recordid, ptmrn, ptlastname
         , priorarrivaldate, priortreatmentdate, firstfhccarrivaldate, patient_list_rectime)
 
# Have staff review any of these where treatment date is after first FHCC arrival, but prior Rx arrival date is not
late_outside_treatment <- allprior |>
  mutate(  priorarrivaldate    = arrivaldate
         , priortreatmentdate  = treatmentdate
         , afterfirstarrival   = priorarrivaldate   > firstfhccarrivaldate
         , afterfirsttreatment = priortreatmentdate > firstfhccarrivaldate 
         , reviewreason = paste('Outside arrival on',priorarrivaldate,
                                'is before first arrival on',firstfhccarrivaldate,
                                'but associated rx on',priortreatmentdate,'is not.')) |>
  filter(  rxline!="" & afterfirsttreatment  & !afterfirstarrival) |>
  select(  recordid, ptmrn, ptlastname, reviewreason, rxline
         , priorarrivaldate, priortreatmentdate, firstfhccarrivaldate, patient_list_rectime)

# Discrepancy report
outside_rx_discreps <- bind_rows(late_outside_arrival, late_outside_treatment)
rm(list = ls(pattern = "^late_outside_"))

if (nrow(outside_rx_discreps) > 0){
  # Create excel for data cleaning
  make_sheet(outside_rx_discreps, scrapdir, 'errordata', 'outside_after_FHCC_arrival', 'xlsx', overwrite=TRUE )
  # Add to table of things to correct
  addrow('todo_df', scrapdir, 'outside_after_FHCC_arrival.xlsx', 'Review for arrival date, treatment date or treatment location errors.')
  cat('ERROR CHECK OUTSIDE treatments dated before the first arrival recorded at FHCC')
} else {
  cat('NO OUTSIDE treatments dated before the first arrival recorded at FHCC')
}

cat('Done\n')

```

## Merge prior arrivals with FHCC arrivals

These records represent all treatment attempts we know about for each patient. Some of these are for treatment that happened outside of our institution, those have an arrivaltable of "prearrival"

```{r eval=FALSE, include=FALSE}
# Merge together all arrival from FHCC and all prior to arrival at FHCC
allarrival_allprior <- bind_rows(allarrival, allprior)  |> 
  arrange(recordid, arrivaldate, treatmentdate, arrivalnum, arrivaltable ) |>
  mutate(row_number = row_number())

cat('Merge prior arrivals with FHCC arrivals')
```

## Rank Patient Arrivals with prev/next

Order the treatment attempts within patient by the chronological order in which they occurred, and chain them to a previous or next arrival for a treatment attempt where possible. The prev & next variables can be handy when comparing between any treatment attempt and what may have happened before or after, particularly nice for calculation of days between events.

```{r eval=FALSE, include=FALSE}
# Adding prev/next arrival information}
# Ranking arrivals
allarrival_allprior <- allarrival_allprior |>
  # mutate(arrival_old = arrival) |>
  arrange(recordid, ptmrn, pmax(arrivaldate, treatmentdate, na.rm = TRUE)) |>
    group_by(recordid) |>
      mutate(
        prevarrivaldate   = lag(arrivaldate),
        prevarrivaltype   = lag(arrivaltype),
        prevarrivalreason = lag(arrivalreason),
        prevarrivalmorph  = lag(arrivalmorph),
        prevarrivalflow   = lead(arrivalflow),
        
        nextarrivaldate   = lead(arrivaldate),
        nextarrivaltype   = lead(arrivaltype),
        nextarrivalreason = lead(arrivalreason),
        nextarrivalmorph  = lead(arrivalmorph),
        nextarrivalflow   = lead(arrivalflow),
        
        # Adding ranking
        arrival = row_number(),
        
        # Count all treatment plans
        allarrivalcnt = n(),

        # Count FHCC arrivals
        fhccarrivalcnt = sum(grepl("^arr", arrivaltable))
        
        
      ) |>
    ungroup() |>
    select(recordid, ptmrn, arrival, everything())

cat('Rank Patient Arrivals with prev/next')
```

## DUPLICATE CHECK

There are a few records that have arrival dates in the induction data that are very close to FH REDCap arrival dates. Most of those were consults at FH, so I filtered consults out of the comparison since it is reasonable to have a FH consult and then return to an outside location for treatment.

I also filtered out arrivals that are not associated with treatment.

```{r eval=FALSE, include=FALSE}
# duplicate check
allarrival_ <- allarrival |> 
  filter(str_detect(rxline, "(?i)consult", negate=TRUE)) |>
  filter(category != 'No Treatment') |>
  select(any_of(idlist), rxline, treatment, everything())

allarrival_dups <- allarrival_ |>
  group_by(recordid) |>
  arrange(arrivaldate) |>
  mutate(next_arrival = lead(arrivaldate)) |>
  mutate(next_category = lead(category)) |>
  mutate(next_rxline = lead(rxline)) |>
  mutate(next_firstinduction = lead(firstinduction)) |>
  mutate(next_arrivaltable = lead(arrivaltable)) |>
  mutate(date_diff = as.numeric(next_arrival - arrivaldate)) |>
  filter(!is.na(date_diff) & abs(date_diff) <= 14) |>
  mutate(arrivalerror = paste0(
    format(arrivaldate, "%Y-%m-%d"), " | ",
    format(next_arrival, "%Y-%m-%d"), 
    " (", date_diff, " days apart)"
  )) |>
  select(recordid, date_diff, arrivalerror, 
         arrivaldate, next_arrival, 
         category, next_category,
         rxline, next_rxline,
         next_firstinduction, firstinduction,
         next_arrivaltable, arrivaltable)

allarrival <- allarrival |>
  left_join(allarrival_dups |> select(recordid, arrivalerror), by = "recordid") |>
  select(any_of(idlist), arrivalerror, everything())
```

