---
title: "Arrival_Cleaning"
author: "Carole Shaw"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

## Purpose

This script is designed to put arrival data together and to distinguish data collected for treatment prior to FHCC arrival, arrivals associated with FHCC care, patients who have not had any arrival data entered yet, and arrivals that not associated with a currative treatment. The plan is to add resulting data from this routine to SQL Server updating it with the weekly workflow so that the data will be consistently compiled and updated over time.

The data created here are joined with other source data, such as transplant data, epic report data, and eventually with CORTEX downloads during the creation of the frozen data.

## Background:

Currently patient treatment plans each generate a new arrival. Usually those are entered in sequential order, and for those that have had their first treatment here the arrival 1 form contains their first treatment here for AML, either frank disease or 10+% blasts. On those records the arrival table is arrival1, the arrivalnum is also 1, and the treatment location is FHCC.

Some patients first were treated at an outside location. There are a few ways that an outside treatment may have been entered into our system. 1) Induction may have been entered into the Diagnosis REDCap, which is now defunct, but induction associated was imported into the 'induction' form of the Estey List REDCap project. 2) An arrival may have been entered as arrivalnum "-1" to indicate an arrival that occurred prior to arrival at FHCC, or 3) an arrival may be entered with location of treatment "OUT" to indicate treatment from an outside location, often (but not always), indicating treatment prior to arrival at FHCC.

We formerly did not enter outside treatment into arrival forms, when that was the case location was implied to be FHCC. The default for location is therefore FHCC when no location is given.

There are some cases where an outside arrival is dated *after* the first FHCC arrival. When that is not associated with hospice/palliative/no treatment those cases should be set aside for review.

If corrections are made to data in REDCap, the REDCap data need to be imported anew for corrections to be reflected here.

## Plan

So the logic in this program is:

1)  Identify any treatment imported from the defunct Diagnosis REDCap into the induction table (**done**)

2)  Merge arrivals identified from the induction table with those from REDCap arrival{n} forms (**done**)

3)  From the allarrival table, find the earliest/last arrival for each patient into tables first(last)\_fhcc_arrival (**done**)

4)  Assign sequence to patient arrivals ordering by arrivaldate, treatmentdate. Using treatment date because many of the outside treatments don't have a concept of arrival. (**done**)

5)  Check for information about the treatment RGNumber associated with the protocol listed from REDCap, mainly for numeric treatment descriptions but looking at the treatment or protocol matching protocol_lkup columns with aliases for the protocol in the column IRB.

6)  Add pointers to each arrival to the previous arrival and the next arrival if such arrivals exist. Pointers currently show the next and previous: category, arrival, arrivaldate, treatmentdate, treatment, and rxline. (**done**)

7)  Copy the allarrival data to SQL Server. (**done**)

8)  Add this script to the weekly workflow, just after import of REDCap data. (**done**)

9)  Run the script: Coding and Data Acquisition/R Scripts/Frozen to SQLServer/**Frozen_Data_to_SQLServer.Rmd** to update the frozen data. The script exists, and has been retooled to expect an allarrival table on SQL Server at frozen.allarrival. (**done**)

# SETUP

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, fig.show = 'asis')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

#### LIBRARIES AND FUNCTIONS

Most libraries are loaded with icicles. Recompile the icicle functions every time to incorporate changes.

```{r library}
# rm(crystals)
# codedir <<- gsub("\\\\", "/", Sys.getenv('R_CODE', unset = NA))
# oldcodedir <<- gsub("\\\\", "/", Sys.getenv('FROZEN_PATH', unset = NA))
srcpath <- normalizePath(file.path(Sys.getenv("R_CODE"), "AML-Libraries", "icicle_setup.R"))
# C:/Users/cmshaw/R Project/AML-Libraries/icicle_setup.R
source(srcpath)
function_rmds = c(
      "icicle_functions.Rmd",
      "sql_functions.Rmd",
      "import_export.Rmd",
      "redcap_api.Rmd",
      "karyotype_function_set.Rmd",
      "timelink.Rmd")
icicle_setup(function_rmds = function_rmds)
idlist      <- c('recordid', 'ptmrn', 'ptlastname')
source("R/Intensity Standardization Code.R")
```

```{r helper}
# summarize_epic_treatment()
# this helper is used in Import EPIC Reports, so grabbing it from there
rmd_path <- "helper.Rmd"
source_cache(file_path = rmd_path, source = TRUE) # source=TRUE is the default, here for clarity
```

## SQL SERVER DATA

Load Some Data from SQL Server

```{r}
# columns of interest
patient_list_cols <- c("recordid",         "ptmrn",        "ptlastname", "ptbirthdate", "dxdate",    
                       "dxmorphdate",      "dxmorphsrc",   "dxmorph",    "dxmorphnum",  "dxmorphcell",
                       "dxflowdate",       "dxflowsrc",    "dxflow",     "dxflownum",
                       "uploaded_to_ctms", "uploaded_to_ctms_date",      "firstarrival", "lastarrival",
                       "firstinduction",   "patient_list_rectime",       "primary_subject_id")

specialty_note_cols <- c("recordid", "ptmrn", "ptlastname",     
                         "ptdeathdate", "ptlastchecked",
                         "lastknownalive", "lastknownremission", "lastknownactivedisease",   
                         "specialty_note_rectime")

local_ <- .pkg_env
get_rc_dictionary('esteylist', ddname = "rcdd_exp", reload = TRUE, explode = TRUE)
local_$rcdd_exp = rcdd_exp
```

```{r eval=FALSE, include=FALSE}
# tables of interest
patient_list <- sql_get_table('esteylist_label.patient_list')
patient_list <- sql_get_table('esteylist_label.patient_list') |> 
    filter(ptlastname != 'PATIENT') |>
    mutate(recordid = as.character(recordid)) |> 
    select(any_of(patient_list_cols)) |>
    convert_date_cols()

specialty_note <- sql_get_table('esteylist_label.specialty_note') |> 
    filter(ptlastname != 'PATIENT') |>
    mutate(recordid = as.character(recordid)) |>
    select(any_of(specialty_note_cols)) |>
    convert_date_cols()

project_specific_info <- sql_get_table('esteylist.project_specific_info') |> 
    filter(ptlastname != 'PATIENT') |>
    mutate(recordid = as.character(recordid)) |>
    group_rc_chkbox() |>
    select(any_of(idlist), special_population)

```

```{r}
# tables of interestpatient_list <- get_rc_table('esteylist', 'patient_list')
patient_list <- get_rc_table('esteylist', 'patient_list')
specialty_note <- get_rc_table('esteylist', 'specialty_note')
project_specific_info <- get_rc_table('esteylist', 'project_specific_info')
```

## Pre-Arrivals from the Induction Table

The induction table contains information from the Diagnosis REDCap that was ported to the Estey List REDCap instance. These are supposed to be treatments from outside locations, but I believe are not always. Once we sequence the arrivals if there are two that are very close or on the same date, we should exclude one or the other.

```{r}
# previous to first arrival induction information
induction_cols <- c("recordid",
    "ptmrn",
    "ptlastname",
    "arrivaltable",
    "redcap_table",
    "arrival",
    "rxline",
    "treatmentlocation",
    "treatmentdate",
    "treatment",
    "responsedate",
    "response",
    "mrd",
    "relapsedate",
    "relapsedesc")

# checking to assure that the induction table is for outside induction information
# disease_features <- get_rc_table('diagnosis_for_aml_arrivals', 'disease_features')|>
#   group_rc_chkbox() |>
#   convert_date_cols() |>
#   filter(df_inductionlocation == 2)
# df_induction <- get_rc_table('diagnosis_for_aml_arrivals', 'induction') |>
#   group_rc_chkbox() |>
#   convert_date_cols() |>
#   select(diagnosisid, ptmrn, recordid, starts_with('in_cycle1'))
# xx <- inner_join(disease_features, df_induction, by = 'ptmrn') |>
#   filter(!is.na(in_cycle1date))

induction_ <- get_rc_table('esteylist','induction')

# zz <- induction <- induction_ |> group_rc_chkbox()


local_ <- .pkg_env
get_rc_dictexp('esteylist')
local_$rcdd_exp <- rcdd_exp

# tic("group_rc_chkbox_ timing")
# zz <- induction <- induction_ |> group_rc_chkbox_()
# toc()
# 
# tic("group_rc_chkbox timing")
# zz <- induction <- induction_ |> group_rc_chkbox()
# toc()
# 
# tic("group_rc_chkbox_dt timing")
# zz <- induction <- induction_ |> group_rc_chkbox()
# toc()

induction <- induction_ |>
  mutate(recordid         = as.character(recordid)
      , arrivaltable      = "induction"
      , redcap_table      = "induction"
      , orig_arrivaltable = "induction"
      , arrival           = as.integer(-1)
      , rxline            = 'Ind'
      # , arrivaldate       = dxdate1
      , treatmentlocation = ''
      , treatmentdate     = dxtreatmentdate1
      , treatment         = dxtreatment1
      , responsedate      = dxresponsedate1
      , response          = dxresponse1
      , mrd               = dxmrd1
      , relapsedate       = inductionrelapsedate
      , relapse           = inductionrelapsenote
      , treatmentdate     = as.Date(treatmentdate)
    ) |> 
  select(any_of(induction_cols)) |> 
  # removed group_rc_chkbox, there weren't any check box fields
  left_join(patient_list,   by = join_by(recordid, ptmrn, ptlastname)) |>
  left_join(specialty_note, by = join_by(recordid, ptmrn, ptlastname)) |>
  mutate(treatmentlocation = if_else(str_detect(firstinduction, 'First AML induction here'), 'FHCC', 'OUT')) |>
  filter(!is.na(treatmentdate)) |>
  arrange(ptmrn, treatmentdate) |>
  # Apply the custom date conversion function across character columns
  convert_date_cols() |>
  select(recordid, ptmrn, ptlastname, 
         treatmentdate, arrivaltable, redcap_table, arrival, responsecategory=response, everything())

induction <- rx_stand(induction, 'treatment')


print('Done')
```

## ARRIVALS 1-11

The variables for arrivals 1 to 11 are identical with the exception of the suffix or prefix for cycles, i.e. arrivaldate1 vs arrivaldate3, or a1c1treatmentdate vs a3c1treatmentdate.

```{r}
# make sure the exploded dictionary is in scope
local_ <- .pkg_env
get_rc_dictexp('esteylist')
local_$rcdd_exp = rcdd_exp

tblcnt      <- 11

arrival_    <- NULL
notreat_    <- NULL
prior_      <- NULL

allarrival  <- NULL
alloutnorx  <- NULL
allprior    <- NULL
alltreat    <- NULL
allnotreat  <- NULL

index <- n <- 0

# Apply the function to the arrival dataframe
# zeroth table is empty
# arrival_ <- get_rc_table('esteylist', 'arrival1')[0,]
  
# looping through each table arrival1 up to arrival11
for (n in 1:tblcnt-1) {
# for (n in 1:2) {
  
  # designed to allow stepping through easier
  if(index==0) {n=n+1} else {n=index}

  # table for this loop
  tbl <- paste0('arrival',  n)
  arrival_ <- get_rc_table('esteylist', tbl)

  if (nrow(arrival_)==0) {
    cat("Skipping empty arrival ", tbl)
    next
  } else {
    cat(tbl, '\n')
  }

  arrival_ <- arrival_ |>
    
    # remove test patients
    filter(ptlastname != 'PATIENT') |>
    # Apply the custom grouping of REDCap checkbox fields into a single column
    group_rc_chkbox() |>
    mutate(
      # Convert recordid to character type  (not sure if this is needed?)
      recordid = as.character(recordid)
      # Add a row number at the start of the df
      , row_number = row_number() + 10000*n,
      # Add a tablename column
      , arrivaltable = tbl
      , redcap_table = tbl
      , orig_arrivaltable = tbl
    ) |>
    # Apply the custom date conversion function across character columns
    convert_date_cols() |>
    # Apply multiple renames in one `rename_with`
    rename_with(~ {
      col_name <- . # col_name is a generic for all columns to meet the pattern
      # Rename "a{n}" fields to "a_"
      col_name <- gsub(glue("^a{n}"), "a_", col_name)
      # Rename "arrival{n}note" fields specifically to "arrival_note"
      col_name <- gsub(glue("^arrival{n}note$"), "arrival_note", col_name)
      # Rename "arrival{n}_" fields to "arrival_"
      col_name <- gsub(glue("^arrival{n}_"), "arrival_", col_name)
      # Rename "arrival{n}" fields to "arrival_"
      col_name <- gsub(glue("^arrival{n}"), "arrival_", col_name)
      # Remove trailing "{n}" from other fields
      col_name <- gsub(glue("{n}$"), "", col_name)
      # Return the transformed column name
      col_name
    }) |>
    mutate(across(everything(), as.character)) |>
    mutate(  arrivalcategory=NA_character_
             # using regular expressions to flag those that didn't get currative treatment at fhcc
             , currativerx_fhcc = is_currative_rx_fhcc(treatment)) |>
    select(row_number, any_of(idlist)
           , any_of(c('arrivaldate', 'arrival_rectime', 'arrivaltable')), 
           redcap_table, treatment, starts_with('curr'), everything())
  
  cols <- arrival_ |> select(ends_with("date")) |> names()

  arrival_ <- arrival_ |>
    left_join(patient_list,   by = idlist) |>
    left_join(specialty_note, by = idlist) |>
    left_join(project_specific_info, by = idlist) |>
    convert_columns('Date', cols) |>
    # convert_date_cols(skip_cols = c("row_number")) |>
    select(row_number, any_of(idlist), primary_subject_id
           , starts_with('special_population_')
           , any_of(c('arrivalcategory', 'arrival_rectime', 'arrivaltable'))
           , redcap_table, treatment, starts_with('curr'), everything())
  


  # set empty structures in the first loop
  if (n==1) {
    allarrival <- alloutnorx <- allprior <- alltreat <- allnotreat <- as.data.frame(arrival_)[0,]
  }
  
  # includes all attempts including arrival w/o treatment and outside treatment
  allarrival <- bind_rows(allarrival, arrival_)


  # clear out tables from last loop
  rm(list = ls(pattern = "^prior|^arrival|^treat|^combo|^outno|^empty"))

}

```
## JOIN OUTSIDE INDUCTION TO ARRIVALS

```{r}
# facilitate date conversion and exclude these from the conversion process
exclude_regex_list = c('id$', 
                       'mrn$', 
                       'note\\d?$|cycles\\d?$', 
                       'procedure$|calc$', 
                       'observations$', 
                       'summary$', 
                       '^bulk_upload', 
                       'backup$', 
                       'fish$', 
                       'history$', 
                       'comment$', 
                       'category$', 
                       'desc(ription)?$', 
                       'result$', 
                       'name$')
exclude_regex = paste(exclude_regex_list, collapse = '|')

# backups used when debugging
induction_  <- induction
allarrival_ <- allarrival
patient_list_ <- patient_list


psi <- project_specific_info |> select(any_of(idlist), starts_with('special_population_'))

# reset in case I am debugging
patient_list <- patient_list_ # debug copy
allarrival   <- allarrival_   # debug copy
induction    <- induction_    # debug copy


patient_list <- patient_list |> convert_date_cols(exclude_regex = exclude_regex)
allarrival   <- allarrival   |> convert_date_cols(exclude_regex = exclude_regex)

allpreabstract <- anti_join(patient_list, allarrival) |>
  left_join(psi) |>
  mutate(category = 'Patient has not been abstracted in REDCap',
         arrivaltable = 'patient_list',
         redcap_table = 'patient_list') |>
  filter( recordid > 0 ) |>
  filter(toupper(ptlastname) != "PATIENT") |>
  select(uploaded_to_ctms, everything()) |>
  filter(uploaded_to_ctms == '1')



# place the patients who have not yet been abstracted into the list of all arrivals
allarrival   <- bind_rows(allarrival, allpreabstract) |>
  mutate(across(
           .cols = names(allarrival),
           .fns  = ~ sql_convert_type(.x, alltreat[[cur_column()]])),
         category = coalesce(category, 'Arrival 1 abstracted')) |>
  # convert_date_cols(exclude_regex=exclude_regex) |>
  select(row_number, 
         any_of(c(idlist, 'category', 'arrivaldate', 'arrival_rectime', 'arrivaltable'
                , 'redcap_table', 'treatment', 'currativerx_fhcc'))
                , everything())

neatfreq(allarrival$category)

# Align date-column types across the two frames before the row-bind.
# Root cause: `allarrival` carries arrival/treatment dates as class Date while the
# `induction` frame (built from the ported Diagnosis REDCap) still holds some of
# the same columns as character. bind_rows() then either errors ("Can't combine
# <date> and <character>") or, once one side is coerced with a bare as.Date(),
# silently turns non-ISO strings into NA. Previously only columns whose Date-ness
# *differed* were coerced (compare_col_types) and with as.Date()'s default format.
# Fix: coerce every column that is Date in EITHER frame (and common to both) on
# BOTH sides using as_date_safe(), which tries known formats and reports parse
# failures instead of dropping them silently.
induction  <- induction_ # reset in case I am debugging

date_cols <- union(
  names(allarrival)[vapply(allarrival, inherits, logical(1), "Date")],
  names(induction )[vapply(induction , inherits, logical(1), "Date")]
)
date_cols  <- intersect(date_cols, intersect(names(allarrival), names(induction)))
allarrival <- allarrival |> mutate(across(any_of(date_cols), as_date_safe))
induction  <- induction  |> mutate(across(any_of(date_cols), as_date_safe))

allarrival <- bind_rows(allarrival, induction |> mutate(category='Prior Treatment Outside')) |>
  mutate(row_number=row_number()+10000,
    currativerx_any  = coalesce(currativerx_fhcc, is.na(currativerx_fhcc) & arrivaltable=="induction")) |>
  arrange(ptmrn, arrivaldate)

dups <- allarrival |>
  mutate(order = if_else(is.na(arrivaldate), as.integer(arrivalnum), as.integer(arrivaldate))) |>
  arrange(ptmrn, order) |>
  group_by(ptmrn) |>
  mutate(
    dup_arrivaldate = duplicated(order) |
                      duplicated(order, fromLast = TRUE)
  ) |>
  ungroup() |>
  select(recordid, ptmrn, ptlastname, arrivalnum, order, arrivaldate, dup_arrivaldate)

# delete dated copies from old runs
rm(list=ls(pattern = '^allarrival_202'))

# make a dated backup copy in case we are debugging
backup_name.1 <- paste0('allarrival_', as.character(now()))
assign(backup_name.1, allarrival)
allarrival <- get(backup_name.1)
```

```{r eval=FALSE, include=FALSE}
# make sure the exploded dictionary is in scope
local_ <- .pkg_env
get_rc_dictexp('esteylist')
local_$rcdd_exp = rcdd_exp

tblcnt      <- 11

arrival_    <- NULL
notreat_    <- NULL
prior_      <- NULL

allarrival  <- NULL
alloutnorx  <- NULL
allprior    <- NULL
alltreat    <- NULL
allnotreat  <- NULL

index <- n <- 0

# Apply the function to the arrival dataframe
# zeroth table is empty
arrival_ <- get_rc_table('esteylist', 'arrival1')[0,]
  
# looping through each table arrival1 up to arrival11
for (n in 1:tblcnt-1) {
# for (n in 1:2) {
  
  # designed to allow stepping through easier
  if(index==0) {n=n+1} else {n=index}

  # table for this loop
  tbl <- paste0('arrival',  n)
  arrival_ <- get_rc_table('esteylist', tbl)

  if (nrow(arrival_)==0) {
    cat("Skipping empty arrival ", tbl)
    next
  } else {
    cat(tbl, '\n')
  }

  arrival_ <- arrival_ |>
    
    # remove test patients
    filter(ptlastname != 'PATIENT') |>
    # Apply the custom grouping of REDCap checkbox fields into a single column
    group_rc_chkbox() |>
    mutate(
      # Convert recordid to character type  (not sure if this is needed?)
      recordid = as.character(recordid)
      # Add a row number at the start of the df
      , row_number = row_number() + 10000*n,
      # Add a tablename column
      , arrivaltable = tbl
      , redcap_table = tbl
      , orig_arrivaltable = tbl
    ) |>
    # Apply the custom date conversion function across character columns
    convert_date_cols() |>
    # Apply multiple renames in one `rename_with`
    rename_with(~ {
      col_name <- . # col_name is a generic for all columns to meet the pattern
      # Rename "a{n}" fields to "a_"
      col_name <- gsub(glue("^a{n}"), "a_", col_name)
      # Rename "arrival{n}note" fields specifically to "arrival_note"
      col_name <- gsub(glue("^arrival{n}note$"), "arrival_note", col_name)
      # Rename "arrival{n}_" fields to "arrival_"
      col_name <- gsub(glue("^arrival{n}_"), "arrival_", col_name)
      # Rename "arrival{n}" fields to "arrival_"
      col_name <- gsub(glue("^arrival{n}"), "arrival_", col_name)
      # Remove trailing "{n}" from other fields
      col_name <- gsub(glue("{n}$"), "", col_name)
      # Return the transformed column name
      col_name
    }) |>
    mutate(across(everything(), as.character)) |>
    mutate(  arrivalcategory=NA_character_
             # using regular expressions to flag those that didn't get currative treatment at fhcc
             , currativerx_fhcc = is_currative_rx_fhcc(treatment)) |>
    select(row_number, any_of(idlist)
           , any_of(c('arrivaldate', 'arrival_rectime', 'arrivaltable')), 
           redcap_table, treatment, starts_with('curr'), everything())
  
  cols <- arrival_ |> select(ends_with("date")) |> names()

  arrival_ <- arrival_ |>
    left_join(patient_list,   by = idlist) |>
    left_join(specialty_note, by = idlist) |>
    left_join(project_specific_info, by = idlist) |>
    convert_columns('Date', cols) |>
    # convert_date_cols(skip_cols = c("row_number")) |>
    select(row_number, any_of(idlist), primary_subject_id
           , starts_with('special_population_')
           , any_of(c('arrivalcategory', 'arrival_rectime', 'arrivaltable'))
           , redcap_table, treatment, starts_with('curr'), everything())
  
  # # outside and not currative
  # outnorx_ <- arrival_ |>
  #   filter(!currativerx_fhcc) |>
  #   filter(a_treatmentlocation == 'OUT' | a_c1treatmentlocation == 'OUT') |>
  #   mutate(arrivalcategory = 'outnotreat')
  # 
  # # identify rows not associated with a treatment at FHCC
  # # notreat_ <- arrival_ |> filter(str_detect(treatment, '(?i)pice|pall|consult|opinion|surv|2nd\\s|(no)+t?\\streat'))
  # notreat_ <- arrival_ |> filter(!currativerx_fhcc) |> mutate(arrivalcategory='notreat')
  # 
  # # keep only rows where there was treatment at FHCC
  # treat_  <- arrival_ |> filter(currativerx_fhcc) |> mutate(arrivalcategory='treated')
  # 
  # # remove prior treatments from the arrival data -- these are outside treatments that were missed above
  # prior_ <- bind_rows(
  #   arrival_ |> filter(rxline == 'Treatment prior to AML induction'),
  #   arrival_ |> filter(a_c1treatmentlocation == 'OUT' & a_treatmentlocation == ''),
  #   arrival_ |> filter(a_treatmentlocation == 'OUT')
  # ) |>
  #   distinct() |>
  #   mutate(arrivalcategory = 'prearrival', arrivaltable = 'prearrival', redcap_table=tbl)
  # 
  # # Remove prior treatment data from the treated data
  # treat_ <- anti_join(treat_, prior_, by=join_by(row_number))
  # 
  # # Remove outside arrivals where no treatment occurred from prior and no treatment
  # prior_ <- anti_join(prior_, outnorx_, by = join_by(row_number))
  # notreat_ <- anti_join(notreat_, outnorx_, by = join_by(row_number))

  # set empty structures in the first loop
  if (n==1) {
    allarrival <- alloutnorx <- allprior <- alltreat <- allnotreat <- as.data.frame(arrival_)[0,]
  }
  
  # includes all attempts including arrival w/o treatment and outside treatment
  allarrival <- bind_rows(allarrival, arrival_)
  # the sum of these should match the all arrival records
  # alltreat   <- bind_rows(alltreat,   treat_)   # arrivals with treatment
  # allprior   <- bind_rows(allprior,   prior_)   # outside care prior to first arrival here
  # alloutnorx <- bind_rows(alloutnorx, outnorx_) # outside care, no treatment
  # allnotreat <- bind_rows(allnotreat, notreat_) # arrivals here not associated with treatment

  # clear out tables from last loop
  rm(list = ls(pattern = "^prior|^arrival|^treat|^combo|^outno|^empty"))
  
  # # Next set
  # tbl = paste0('arrival',  n+1)  
  # arrival_ <- get_rc_table('esteylist', tbl)
  
}

# # this just puts the columns in order
# alloutnorx <- alloutnorx |>
#   mutate(category = 'alloutnorx', 
#          across(
#           .cols = intersect(names(allarrival), names(allprior)),
#           .fns  = ~ sql_convert_type(.x, alltreat[[cur_column()]]))) |>
#   convert_date_cols() |>
#   select(row_number, 
#          any_of(c(idlist, 'category', 'arrivaldate', 'arrival_rectime', 'arrivaltable'
#                 , 'redcap_table', 'treatment', 'currativerx_fhcc'))
#                 , everything())
# 
# allprior <- allprior |>
#   mutate(category = 'allprior', 
#          across(
#           .cols = intersect(names(allarrival), names(allprior)),
#           .fns  = ~ sql_convert_type(.x, alltreat[[cur_column()]]))) |>
#   convert_date_cols() |>
#   select(row_number, 
#          any_of(c(idlist, 'category', 'arrivaldate', 'arrival_rectime', 'arrivaltable'
#                 , 'redcap_table', 'treatment', 'currativerx_fhcc'))
#                 , everything())
# allprior <- bind_rows(allprior, induction) |> mutate(category = 'allprior')
# 
# alltreat <- alltreat |>
#   mutate(category = 'alltreat', 
#          across(
#           .cols = intersect(names(allarrival), names(alltreat)),
#           .fns  = ~ sql_convert_type(.x, alltreat[[cur_column()]]))) |>
#   convert_date_cols() |>
#   select(row_number, 
#          any_of(c(idlist, 'category', 'arrivaldate', 'arrival_rectime', 'arrivaltable'
#                 , 'redcap_table', 'treatment', 'currativerx_fhcc'))
#                 , everything())
# allnotreat <- allnotreat |>
#   mutate(category = 'allnotreat', 
#          across(.cols = intersect(names(allarrival), names(allnotreat)),
#                 .fns  = ~ sql_convert_type(.x, allnotreat[[cur_column()]]))) |>
#   convert_date_cols() |>
#   select(row_number, 
#          any_of(c(idlist, 'category', 'arrivaldate', 'arrival_rectime', 'arrivaltable'
#                 , 'redcap_table', 'treatment', 'currativerx_fhcc'))
#                 , everything())

psi <- project_specific_info |> select(any_of(idlist), starts_with('special_population_'))

allpreabstract <- anti_join(patient_list, allarrival) |>
  left_join(psi) |>
  mutate(category = 'Patient has not been abstracted in REDCap',
         special_population = coalesce(special_population, ''),
         arrivaltable = 'patient_list',
         redcap_table = 'patient_list') |>
  filter(str_detect(special_population, '(?i)^older aml|mucor', negate=TRUE)) |>
  filter( recordid > 0
        & str_detect(ptlastname, "(?i)patient", negate=TRUE)
        & uploaded_to_ctms == "Yes")


allarrival   <- bind_rows(allarrival, allpreabstract) |>
  mutate(across(
           .cols = names(allarrival),
           .fns  = ~ sql_convert_type(.x, alltreat[[cur_column()]]))) |>
  mutate(category = case_when(
      row_number %in% alloutnorx$row_number ~ "Outside Arrival without Treatment",
      row_number %in% allprior$row_number   ~ "Prior Treatment Outside",
      row_number %in% allnotreat$row_number ~ "No Treatment",
      row_number %in% alltreat$row_number   ~ "FH Treatment",
      row_number %in% allpreabstract        ~ "Not yet abstracted",
      TRUE ~ category  # retain existing value if no match
    )
  ) |>
  convert_date_cols() |>
  select(row_number, 
         any_of(c(idlist, 'category', 'arrivaldate', 'arrival_rectime', 'arrivaltable'
                , 'redcap_table', 'treatment', 'currativerx_fhcc'))
                , everything())

# Align date-column types across both frames before the row-bind (see the
# earlier bind for the full root-cause note): coerce every column that is Date
# in EITHER frame (and common to both) on BOTH sides with as_date_safe(), which
# tries known formats and surfaces parse failures instead of silent NA.
date_cols <- union(
  names(allarrival)[vapply(allarrival, inherits, logical(1), "Date")],
  names(induction )[vapply(induction , inherits, logical(1), "Date")]
)
date_cols  <- intersect(date_cols, intersect(names(allarrival), names(induction)))
allarrival <- allarrival |> mutate(across(any_of(date_cols), as_date_safe))
induction  <- induction  |> mutate(across(any_of(date_cols), as_date_safe))

allarrival <- bind_rows(allarrival, induction |> mutate(category='Prior Treatment Outside')) |>
  mutate(row_number=row_number()+10000,
    currativerx_any  = coalesce(currativerx_fhcc, is.na(currativerx_fhcc) & arrivaltable=="induction"))

rslt <- nrow(allarrival) == nrow(alloutnorx) + nrow(alltreat) + nrow(allprior) + nrow(allnotreat) + nrow(allpreabstract)
cat("All records accounted for?", rslt)

rm(list=ls(pattern = '^allarrival_2025'))

backup_name.1 <- paste0('allarrival_', as.character(now()))
assign(backup_name.1, allarrival)
allarrival <- get(backup_name.1)
```


## CATEGORIZE INTENSITY, BACKBONE, TKI's

```{r}
zz<- get(backup_name.1)
rm(allarrival)
allarrival <- get(backup_name.1) |> 
  filter(str_detect(arrivaltable, "patient_list", negate=TRUE)) |>
  mutate(arrival_num = case_when(
      arrivaltable == "induction" ~ 0L,
      str_detect(arrivaltable, "\\d+") ~ parse_integer(str_extract(arrivaltable, "\\d+")),
      TRUE ~ NA_integer_)) |>
  select(any_of(idlist), treatment, -ends_with('_calc'), arrivaltable, arrival_num, everything())

# BACKBONE
allarrival  <- add_backbone(allarrival, treatment, ties = "all")
# INTENSITY
allarrival <- classify_intensity(allarrival, treatment, output = "label")
# MARK TKI'S
allarrival <- tag_tkis(allarrival)
# Arrange so that we can review the pertinent variables
allarrival <- allarrival |> select(any_of(idlist), ends_with(c('calc','tki')), everything())

treatment_detail <- allarrival |>
  filter(arrivalnum == 1, !is.na(treatment)) |>
  select(any_of(idlist), treatment,
         treatmentbackbone_calc, treatmentintensity_calc) |>
  group_by(
    treatmentbackbone_calc,
    treatmentintensity_calc
  ) |>
  summarise(
    n = n(),
    treatment_type_summary = paste(unique(treatment), collapse = "\n"),
    .groups = "drop"
  ) |>
  mutate(
    pct = 100 * n / sum(n),
    pct = sprintf("%.2f", pct)   # format to two decimal places
  ) |>
  arrange(desc(n))

treatment_total <- treatment_detail |>
  summarise(
    treatmentbackbone_calc   = "All Rx Combinations",
    treatmentintensity_calc  = "All Intensities",
    n                        = sum(as.numeric(n)),
    treatment_type_summary   = "All Rx Combinations",
    pct                      = sprintf("%.2f", 100)
  )

treatment_summary <- bind_rows(treatment_detail, treatment_total)

make_sheet(treatment_summary, scrapdir, filename="intensity and backbone mapping summary")

# All of the data have been gathered at this point, but all arrival is arranged
# with one row per arrival, we need to go wide to accommodate the multiple
# arrival forms in REDCap.

# WIDE BACKBONE
backbone_wide <- allarrival |>
  filter(!is.na(arrival_num)) |>
  distinct(across(all_of(idlist)), arrival_num, treatmentbackbone_calc, .keep_all = FALSE) |> 
  tidyr::pivot_wider(
    names_from   = arrival_num,
    values_from  = treatmentbackbone_calc,
    names_prefix = "treatmentbackbone_calc",
    values_fn    = ~ dplyr::first(stats::na.omit(.)),
    values_fill  = NA
  )

# WIDE INTENSITY
intensity_wide <- allarrival |>
  filter(!is.na(arrival_num)) |>
  distinct(across(all_of(idlist)), arrival_num, treatmentintensity_calc, .keep_all = FALSE) |> 
  tidyr::pivot_wider(
    names_from   = arrival_num,
    values_from  = treatmentintensity_calc,
    names_prefix = "treatmentintensity_calc",
    values_fn    = ~ dplyr::first(stats::na.omit(.)),
    values_fill  = NA
  )

# WIDE BCR::ABL TKI
bcrabltki_wide <- allarrival |>
  filter(!is.na(arrival_num) & !is.na(treatmentbcrabl_tki)) |>
  distinct(across(all_of(idlist)), arrival_num, treatmentbcrabl_tki, .keep_all = FALSE) |> 
  tidyr::pivot_wider(
    names_from   = arrival_num,
    values_from  = treatmentbcrabl_tki,
    names_prefix = "treatmentbcrabl_tki",
    values_fn    = ~ dplyr::first(stats::na.omit(.)),
    values_fill  = NA
  )

# WIDE FLT3 TKI
flt3tki_wide <- allarrival |>
  filter(!is.na(arrival_num) & !is.na(treatmentflt3_tki)) |>
  distinct(across(all_of(idlist)), arrival_num, treatmentflt3_tki, .keep_all = FALSE) |> 
  tidyr::pivot_wider(
    names_from   = arrival_num,
    values_from  = treatmentflt3_tki,
    names_prefix = "treatmentflt3_tki",
    values_fn    = ~ dplyr::first(stats::na.omit(.)),
    values_fill  = NA
  )

# COMBINE WIDE DATA
classified_treatment <- intensity_wide |>
  full_join(backbone_wide, by=idlist) |>
  full_join(bcrabltki_wide, by=idlist) |>
  full_join(flt3tki_wide, by=idlist)

rc_classified_treatment <- get_rc_report('esteylist', 386188)

# columns to compare, excluding the ID fields if you want
common_cols <- intersect(names(classified_treatment), names(rc_classified_treatment))
compare_cols <- setdiff(common_cols, idlist)

new_old_classified_treatment <- left_join(classified_treatment |> select(any_of(common_cols)), 
                 rc_classified_treatment  |> select(any_of(common_cols)), 
                 by = idlist, suffix = c("_new", "_old"))

update_classified_treatment <- new_old_classified_treatment %>%
  rowwise() %>%
  mutate(
    update_flag = any(
      c_across(ends_with("_new")) %in% c("", NA) == FALSE &
      c_across(ends_with("_old")) %in% c("", NA)
    )
  ) %>%
  ungroup() %>%
  filter(update_flag) %>%
  select(recordid, ptmrn, ptlastname, ends_with("new")) %>%
  mutate(
    across(
      .cols = ends_with("_new"),
      .fns  = as.factor
    )
  ) %>%
  rename_with(~ str_remove(.x, "_new"), ends_with("_new"))

# QC Check, one record per patient/recordid
length(unique(update_classified_treatment$recordid)) == nrow(update_classified_treatment)

# Output to import directory
make_sheet(update_classified_treatment, rcimportdir, 
           filename = 'classified_treatment_import', 
           time=TRUE, 
           overwrite = TRUE, 
           null_as_blank = TRUE)

join_by <- c(idlist, 'row_number')

allarrival <- allarrival |> select(any_of(join_by), ends_with(c('calc','tki')))

allarrival <- get(backup_name.1) %>%
  left_join(allarrival, by=join_by, suffix = c(".x", "")) |>
  select(any_of(names(allarrival)), treatment, arrivaltable, category, everything(), -ends_with('.x')) |>
  mutate(category=if_else(arrivaltable == 'patient_list', 'Abstraction of Arrivals not Complete', category))

backup_name.2 <- paste0('allarrival_', as.character(now()))
assign(backup_name.2, allarrival)

```

### RUN PART TWO?


```{r}
runall <- coalesce(.pkg_env$runall, TRUE)
if (runall) source_cache('C:/Users/cmshaw/R Project/Arrival_Cleaning/Arrival_Cleaning Part2.Rmd', source=TRUE)
```