---
title: "Arrival_Cleaning"
author: "Carole Shaw"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

## Purpose

This script is designed to put arrival data together and to distinguish data collected for treatment prior to FHCC arrival, arrivals associated with FHCC care, patients who have not had any arrival data entered yet, and arrivals that not associated with a curative treatment at FHCC (i.e. comfort care/consult) . The plan is to add resulting data from this routine to SQL Server updating it with the weekly workflow so that the data will be consistently compiled and updated over time.\
\
The data created here are joined with other source data, such as transplant data, epic report data, and eventually with CORTEX downloads during the creation of the frozen data.

## Background:

Currently patient treatment plans each generate a new arrival. Usually those are entered in sequential order, and for those that have had their first treatment here the arrival 1 form contains their first treatment here for AML, either frank disease or \>10% blasts. On those records the arrival table is arrival1, the arrivalnum is also 1, and the treatment location is FHCC.

Some patients first were treated at an outside location. There are a few ways that an outside treatment may have been entered into our system. 1) Induction may have been entered into the Diagnosis REDCap, which is now defunct, but induction associated was imported into the 'induction' form of the Estey List REDCap project. 2) An arrival may have been entered as arrivalnum "-1" to indicate an arrival that occurred prior to arrival at FHCC, or 3) an arrival may be entered with location of treatment "OUT" to indicate treatment from an outside location, often (but not always), indicating treatment prior to arrival at FHCC.

We formerly did not enter outside treatment into arrival forms, when that was the case location was implied to be FHCC. The default for location is therefore FHCC when no location is given.

There are some cases where an outside arrival is dated *after* the first FHCC arrival. When that is not associated with hospice/palliative/no treatment those cases should be set aside for review.

If corrections are made to data in REDCap, the REDCap data need to be imported anew for corrections to be reflected here.

## Plan

So the logic in this program is:

1)  Identify any treatment imported from the defunct Diagnosis REDCap into the induction table (**done**)
2)  Separate arrivals outside treatments into allprior table (**done**)
3)  Separate arrivals represent a treatment at FHCC or without a treatment location into allarrival table (**done**)
4)  Merge arrivals identified from the induction table with those from the allprior, remove duplicates (**done**)
5)  From the allarrival table, find the earliest arrival for each patient into a table first_fhcc_arrival (**done**)
6)  Compare allprior arrivals to the first_fhcc_arrival for each patient to find those that need to be set aside for review, ignoring any where the treatment was No Treatment, Hospice or Palliative. (**done**)
7)  If data cleaning to the REDCap instance occurred at step 6, update REDCap data on SQL server and then re-run steps 1-6, otherwise go on to step 8. (**done**)
8)  Put together the allarrival table with those remaining in the allprior table. (**done**)
9)  Assign sequence to patient arrivals ordering by arrivaldate, treatmentdate. Using treatment date because many of the outside treatments don't have a concept of arrival. (**done**)
10) Copy the allarrival data to SQL Server. (**done**)
11) Add this script to the weekly workflow, just after import of REDCap data. (**done**)
12) Run the script: Coding and Data Acquisition/R Scripts/Frozen to SQLServer/**Frozen_Data_to_SQLServer.Rmd** to update the frozen data. The script exists, and has been retooled to expect an allarrival table on SQL Server at frozen.allarrival. (**done**)

# SETUP

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, fig.show = 'asis')
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

#### LIBRARIES AND FUNCTIONS

Most libraries are loaded with icicles. Recompile the icicle functions every time to incorporate changes.

```{r library}
codedir <<- gsub("\\\\", "/", Sys.getenv('FROZEN_PATH', unset = NA))
source(paste0(codedir,'/icicle_setup.R')); icicle_setup()
```

## SQL SERVER DATA

Load Some Data from SQL Server

```{r}
# columns of interest
patient_list_cols <- c("recordid",         "ptmrn",        "ptlastname", "ptbirthdate", "dxdate",    
                       "dxmorphdate",      "dxmorphsrc",   "dxmorph",    "dxmorphnum",  "dxmorphcell",
                       "dxflowdate",       "dxflowsrc",    "dxflow",     "dxflownum",
                       "uploaded_to_ctms", "uploaded_to_ctms_date",      "firstarrival", "lastarrival",
                       "firstinduction",   "patient_list_rectime",       "primary_subject_id")

specialty_note_cols <- c("recordid", "ptmrn", "ptlastname",     
                         "ptdeathdate", "ptlastchecked",
                         "lastknownalive", "lastknownremission", "lastknownactivedisease",   
                         "specialty_note_rectime")

# tables of interest
patient_list <- sql_get_table('esteylist_label.patient_list') |> 
    filter(ptlastname != 'PATIENT') |>
    mutate(recordid = as.character(recordid)) |> 
    select(any_of(patient_list_cols))

specialty_note <- sql_get_table('esteylist_label.specialty_note') |> 
    filter(ptlastname != 'PATIENT') |>
    mutate(recordid = as.character(recordid)) |>
    select(any_of(specialty_note_cols))

# Make sure the REDCap data dictionary is loaded for the current section's schema
get_rc_dictionary('esteylist')
```

## Pre-Arrivals from the Induction Table

```{r}
# previous to first arrival induction information
induction_cols <- c("recordid",
    "ptmrn",
    "ptlastname",
    "arrivaltable",
    "arrival",
    "rxline",
    "treatmentlocation",
    "treatmentdate",
    "treatment",
    "responsedate",
    "response",
    "mrd",
    "relapsedate",
    "relapsedesc")

induction <- sql_get_table(glue('esteylist_label.induction')) |> 
  # Apply the custom date conversion function across character columns
  convert_date_cols() |>
  # Apply the custom grouping of REDCap checkbox fields into a single column
  group_rc_chkbox() |>
  mutate(recordid         = as.character(recordid)
      , arrivaltable      = "induction"
      , orig_arrivaltable = "induction"
      , arrival           = as.integer(-1)
      , rxline            = 'Ind'
      # , arrivaldate       = dxdate1
      , treatmentlocation = 'OUT'
      , treatmentdate     = dxtreatmentdate1
      , treatment         = dxtreatment1
      , responsedate      = dxresponsedate1
      , response          = dxresponse1
      , mrd               = dxmrd1
      , relapsedate       = inductionrelapsedate
      , relapse           = inductionrelapsenote
      , treatmentdate     = as.Date(treatmentdate)
    ) |> 
  select(any_of(induction_cols)) |> 
  left_join(patient_list,   by = join_by(recordid, ptmrn, ptlastname)) |>
  left_join(specialty_note, by = join_by(recordid, ptmrn, ptlastname)) |>
  filter(!is.na(treatmentdate)) |>
  arrange(ptmrn, treatmentdate) |>
  select(recordid, ptmrn, ptlastname, treatmentdate, arrivaltable, arrival, everything())

print('Done')
```

## ARRIVALS 1-11

```{r}
schema      <- 'frozen'
tblcnt      <- 11

arrival_    <- NULL
notreat_    <- NULL
prior_      <- NULL

allprior    <- NULL
alltreat_   <- NULL
allnotreat  <- NULL


for (n in 1:tblcnt) {
  # table for this loop
  tbl <- paste0('arrival',  n)
  
  # Apply the function to the arrival dataframe
  arrival_ <- sql_get_table(glue('esteylist_label.arrival{n}')) |>
    # remove test patients
    filter(ptlastname != 'PATIENT') |>
    # Apply the custom grouping of REDCap checkbox fields into a single column
    group_rc_chkbox() |>
    mutate(
      # Convert recordid to character type
        recordid = as.character(recordid)
      # Add a row number
      , row_number = row_number(), .before = 1    
      # Add a tablename column
      , arrivaltable = tbl
      , orig_arrivaltable = tbl
    ) |>
    # Apply the custom date conversion function across character columns
    convert_date_cols() |>
    # Apply multiple renames in one `rename_with`
    rename_with(~ {
      col_name <- .
      # Rename "a{n}" fields to "a_"
      col_name <- gsub(glue("^a{n}"), "a_", col_name)
      # Rename "arrival{n}note" fields specifically to "arrival_note"
      col_name <- gsub(glue("^arrival{n}note$"), "arrival_note", col_name)
      # Rename "arrival{n}" fields to "arrival"
      col_name <- gsub(glue("^arrival{n}"), "arrival", col_name)
      # Remove trailing "{n}" from other fields
      col_name <- gsub(glue("{n}$"), "", col_name)
      # Return the transformed column name
      col_name
    }) |>
    mutate(  arrivalcategory=NA_character_
           # using regular expressions to flag those that didn't get curative treatment
           , currativerx   = !grepl("no\\s+rx|no\\s+rec|spice|pall?i|unknown|comfort|consult|outside|support|opinion|surv|2nd\\s|none|not\\selig|not?\\streat", treatment, ignore.case = TRUE)) |>
    select(row_number, recordid, ptmrn, ptlastname, arrivaldate, arrival_rectime, arrivaltable
           , treatment, currativerx, everything())

  arrival_ <- arrival_ |>
    left_join(patient_list, by = join_by(recordid, ptmrn, ptlastname)) |> 
    left_join(specialty_note, by = join_by(recordid, ptmrn, ptlastname)) |>
    select(row_number, recordid, ptmrn, ptlastname, primary_subject_id, arrivalcategory, everything())
  
  # REFACTOR, this comment doesn't make sense:: Retype arrival to match SQL Server, removed code that did this twice
  # Only do for arrivals after 1 since the alltreat_ df does not exist yet
  if (n > 1) {
    arrival_ <- arrival_ |>
      mutate(across(
        .cols = intersect(names(arrival_), names(alltreat_)),
        .fns = ~ sql_convert_type(.x, alltreat_[[cur_column()]])
      ))
  }

  # identify rows not associated with a treatment
  # notreat_ <- arrival_ |> filter(str_detect(treatment, '(?i)pice|pall|consult|opinion|surv|2nd\\s|(no)+t?\\streat'))
  notreat_ <- arrival_ |> filter(!currativerx) |> mutate(arrivalcategory='notreat')

  # keep only rows where there was treatment
  arrival_  <- arrival_ |> filter(currativerx) |> mutate(arrivalcategory='treated')

  # # combine, and make sure to remove dups
  # prior_ <- bind_rows(lapply(ls(pattern = "^prior_\\d+"), get)) # prior_ <- bind_rows(prior_, prior1, prior2, prior3) |> distinct()
  # prior_ <- distinct(prior_) |> mutate(arrivalcategory='prearrival', arrivaltable='prearrival')  
  # 
  
  # remove prior treatments from the arrival data
  prior_1 <- arrival_ |> filter(rxline                == 'Treatment prior to AML induction')
  prior_2 <- arrival_ |> filter(a_c1treatmentlocation == 'OUT' & a_treatmentlocation   == '')
  prior_3 <- arrival_ |> filter(a_treatmentlocation   == 'OUT')
  
  # combine, and make sure to remove dups
  prior_ <- bind_rows(lapply(ls(pattern = "^prior_\\d+"), get)) # prior_ <- bind_rows(prior_, prior1, prior2, prior3) |> distinct()
  prior_ <- distinct(prior_) |> mutate(arrivalcategory='prearrival', arrivaltable='prearrival')


  # Remove prior treatment data from the arrival data
  arrival_ <- anti_join(arrival_, prior_, by=join_by(row_number)) 
  
  # set empty structures if needed, usually in just the first loop, these empty structures should be the same
  # if (is.null(allprior) & is.null(alltreat_)) {
  if (n==1) {
    # empty prior structure
    allprior   <- as.data.frame(prior_)[0,]
    # empty arrival structure
    alltreat_  <- as.data.frame(arrival_)[0,]
    # empty structure for no treatment
    allnotreat <- as.data.frame(notreat_)
  }
  
  
  # Join arrivals -- arrivals with treatment
  alltreat_ <- bind_rows(alltreat_, arrival_) # |> mutate(row_number = row_number())
  
  # Join priors -- care prior to first arrival here
  allprior  <- bind_rows(allprior, prior_) |> mutate(row_number = row_number() + nrow(alltreat_))

  # Join no treatment -- arrivals that were not associated with treatment
  allnotreat <- bind_rows(allnotreat, notreat_) |> mutate(row_number = row_number() + nrow(alltreat_) + nrow(allprior))
  

}

# this just puts the columns in order
allprior   <- allprior |> select(row_number, recordid, ptmrn, ptlastname, arrivaldate, arrival_rectime, arrivaltable
           , treatment, currativerx, everything())
alltreat_  <- alltreat_ |> select(row_number, recordid, ptmrn, ptlastname, arrivaldate, arrival_rectime, arrivaltable
           , treatment, currativerx, everything())
allnotreat <- allnotreat |> select(row_number, recordid, ptmrn, ptlastname, arrivaldate, arrival_rectime, arrivaltable
           , treatment, currativerx, everything())


# clear out tables from last loop
rm(list = ls(pattern = "^prior"))
rm(list = ls(pattern = "^arrival"))
rm(list = ls(pattern = "^notreat"))
```

## SEPARATE INTO 3 ARRIVAL CATEGORIES

#### allprior   - arrivals prior to arrival af fhcc
#### allarrival - arrivals for patients that were treated or are pre-abstraction
#### allnotreat - patients who were seen in clinic or tele-health but did not get treatment

```{r}

# restore if needed:
# Keep a copy, but realize that the data have been incorporated into the df's allarrival, allnotreat, and allprior

## Before binding allprior and induction check for duplicate introduction
allprior_  <- allprior
testbind   <- bind_rows(allprior_, induction) |> distinct(recordid, ptmrn, ptlastname, arrivaldate, arrivaltable, .keep_all = TRUE)
outlier    <- anti_join(allprior_, testbind)
outlierset <- left_join(outlier, allprior_, by=c('recordid', 'ptmrn', 'ptlastname', 'arrivaldate', 'arrivaltable')) |> 
  select(recordid, ptmrn, ptlastname
         , starts_with(c('orig_', 'arrivalcategory', 'arrivaldate', 'arrivaltable', 'treatment.', 'treatmentdate.')))

# These are duplicate rows
if (nrow(outlierset)>0) {
    make_sheet(outlierset, qcdir, filename="multiple_outside_treatment_and_second_opinion_records")
}

# remove duplicate rows
allprior <- anti_join(allprior_, outlier)
# add in induction records
allprior <- bind_rows(allprior, induction) |> 
  mutate(arrivaltable = 'prearrival',
         arrivalcategory = 'prearrival') |>
  arrange(recordid, arrivaldate, treatmentdate, arrivalnum) |>
  # Add a unique arrival identifier as row number
  mutate(row_number = row_number(), .before = 1)

# patients who appear as a patient but do not have an arrival record are presumed to be pre-abstraction
allpreabstract_ <- anti_join(patient_list, bind_rows(alltreat_, allprior, allnotreat)) |>
  distinct(recordid, ptmrn, ptlastname, dxdate, .keep_all = TRUE) |>
  mutate(arrivalcategory = 'preabstract', arrivaltable = 'preabstract')

testbind <- bind_rows(alltreat_, allpreabstract_) %>%
  distinct(recordid, ptmrn, ptlastname, .keep_all = TRUE) %>%
  mutate(row_number = row_number())
outlier  <- testbind %>% count(recordid, ptmrn, ptlastname) %>% filter(n > 1)
# These are duplicate rows
if (nrow(outlier)>0) {
    stop("This should not happen, pats should not be pre-abstraction and have an arrival with treatment")
}

# Keeping the pre-abstraction with the treatment arrivals from alltreat_
allarrival <- bind_rows(alltreat_, allpreabstract_) |> mutate(row_number = row_number())

# Summarize rows, there should be only one row per person per date
outlier <- allarrival |> group_by(recordid, ptmrn, ptlastname, arrivaldate) |>
  mutate(cnt=n()
         , treatmentsummary = paste0(
            ifelse(is.na(treatmentdate), "", format(treatmentdate, '[%m/%d/%Y] ')),
            ifelse(is.na(treatment), "", treatment)),
            sep = " ") |>
  ungroup() |>
  filter(cnt>1) |>
  arrange(recordid, ptmrn, ptlastname, arrivaldate) |>
  select(row_number, recordid, ptmrn, ptlastname, arrivalcategory, arrivaldate, arrivaltable=orig_arrivaltable, treatmentsummary)

if (nrow(outlier)>0) {
    # These are duplicate rows ... multiple arrivals on a single date
    make_sheet(outlier, qcdir, filename="multiple_arrivals_on_a_date")
  }

rm(allprior_)


```

```{r eval=FALSE, message=TRUE, include=FALSE}
rx  <- nrow(alltreat_)
pre <- nrow(allpreabstract_)
cat("All arrival rows        ", nrow(allarrival), glue("\tdf=allarrival (Treated n={rx} and Pre-Abstraction n={pre})"), "\n")
cat("All no treatment rows   ", nrow(allnotreat), "\tdf=allnotreat\n")
cat("All prior arrival rows  ", nrow(allprior),   "\tdf=allprior\n")
df <- freq(allnotreat$treatment)
head(df, 50)
```

## FHCC FIRST/LAST ARRIVAL

```{r}
# identify earliest FHCC arrival
first_fhcc_arrival <- bind_rows(allarrival, allnotreat) |>
  mutate(min_date = pmin(arrivaldate, treatmentdate, na.rm = TRUE)) |>
  select(recordid, min_date) |>
  filter(!is.na(min_date)) |>
  group_by(recordid) |>
  summarise(firstfhccarrivaldate = min(min_date)  # Get min_date for each group
    , .groups = 'drop')

last_fhcc_arrival <- bind_rows(allarrival, allnotreat) |>
  mutate(max_date = pmax(arrivaldate, treatmentdate, na.rm = TRUE)) |>
  select(recordid, max_date) |>
  filter(!is.na(max_date)) |>
  group_by(recordid) |>
  summarise(lastfhccarrivaldate = max(max_date)         # Get max_date for each group            
    , .groups = 'drop')

allarrival <- left_join(allarrival, first_fhcc_arrival) |> left_join(last_fhcc_arrival)
allprior   <- left_join(allprior,   first_fhcc_arrival) |> left_join(last_fhcc_arrival)

```

## PUTTING IT ALL TOGETHER WITH LAG/LEAD VARS

```{r}

all <- bind_rows(allprior, allarrival, allnotreat) |> 
  arrange(recordid, arrivaldate, treatmentdate, arrivalnum, arrivaltable ) |>
  mutate(row = row_number()) |>
  group_by(recordid, ptmrn, ptlastname) |>
  mutate(sequence = row_number(),
         prevarrivaldate=lag(arrivaldate),
         prevarrivalcategory=lag(arrivalcategory),
         prevtreatmentdate=lag(treatmentdate),
         prevtreatment=lag(treatment),
         nextarrivaldate=lead(arrivaldate),
         nextarrivalcategory=lead(arrivalcategory),
         nexttreatmentdate=lead(treatmentdate),
         nexttreatment=lead(treatment)) |>
  ungroup() |>
  select(row, sequence, recordid, ptmrn, ptlastname, primary_subject_id
         , arrivalcategory, firstfhccarrivaldate, lastfhccarrivaldate
         , starts_with(c('arrival', 'prev', 'next')), everything())

```

## ERROR CHECK

### OUTSIDE treatments dated *before* the first arrival recorded at FHCC

Answers the question: Are there arrivals dated after first FHCC arrival/treatment that we have indicated as "OUT" for location?

Then produces an excel document in the scrap directory with records that seem fishy.

```{r}
# Are there prior arrivals dated after first FHCC arrival/treatment

# Have staff review any of these where arrival date is after first FHCC arrival, but Rx date is not
late_outside_arrival <- allprior |>
  mutate(  priorarrivaldate    = arrivaldate
         , priortreatmentdate  = treatmentdate
         , afterfirstarrival   = priorarrivaldate   > firstfhccarrivaldate
         , afterfirsttreatment = priortreatmentdate > firstfhccarrivaldate
         , reviewreason = paste('Arrival date',arrivaldate
                                ,'is after first arrival on',firstfhccarrivaldate)) |>  
  filter(  afterfirstarrival & !afterfirsttreatment) |>
  select(  recordid, ptmrn, ptlastname
         , priorarrivaldate, priortreatmentdate, firstfhccarrivaldate, patient_list_rectime)
 
# Have staff review any of these where treatment date is after first FHCC arrival, but prior Rx arrival date is not
late_outside_treatment <- allprior |>
  mutate(  priorarrivaldate    = arrivaldate
         , priortreatmentdate  = treatmentdate
         , afterfirstarrival   = priorarrivaldate   > firstfhccarrivaldate
         , afterfirsttreatment = priortreatmentdate > firstfhccarrivaldate 
         , reviewreason = paste('Outside arrival on',priorarrivaldate,
                                'is before first arrival on',firstfhccarrivaldate,
                                'but associated rx on',priortreatmentdate,'is not.')) |>
  filter(  rxline!="" & afterfirsttreatment  & !afterfirstarrival) |>
  select(  recordid, ptmrn, ptlastname, reviewreason, rxline
         , priorarrivaldate, priortreatmentdate, firstfhccarrivaldate, patient_list_rectime)

# Discrepancy report
outside_rx_discreps <- bind_rows(late_outside_arrival, late_outside_treatment)
rm(list = ls(pattern = "^late_outside_"))

if (nrow(outside_rx_discreps) > 0){
  # Create excel for data cleaning
  make_sheet(outside_rx_discreps, scrapdir, 'errordata', 'outside_after_FHCC_arrival', 'xlsx', overwrite=TRUE )
  # Add to table of things to correct
  addrow('todo_df', scrapdir, 'outside_after_FHCC_arrival.xlsx', 'Review for arrival date, treatment date or treatment location errors.')
  cat('ERROR CHECK OUTSIDE treatments dated before the first arrival recorded at FHCC')
} else {
  cat('NO OUTSIDE treatments dated before the first arrival recorded at FHCC')
}


```

## Merge prior arrivals with FHCC arrivals

These records represent all treatment attempts we know about for each patient. Some of these are for treatment that happened outside of our institution, those have an arrivaltable of "prearrival"

```{r}
# Merge together all arrival from FHCC and all prior to arrival at FHCC
allarrival_allprior <- bind_rows(allarrival, allprior)  |> 
  arrange(recordid, arrivaldate, treatmentdate, arrivalnum, arrivaltable ) |>
  mutate(row_number = row_number())

cat('Merge prior arrivals with FHCC arrivals')
```

## Rank Patient Arrivals with prev/next

Order the treatment attempts within patient by the chronological order in which they occurred, and chain them to a previous or next arrival for a treatment attempt where possible. The prev & next variables can be handy when comparing between any treatment attempt and what may have happened before or after, particularly nice for calculation of days between events.

```{r}
# Adding prev/next arrival information}
# Ranking arrivals
allarrival_allprior <- allarrival_allprior |>
  # mutate(arrival_old = arrival) |>
  arrange(recordid, ptmrn, pmax(arrivaldate, treatmentdate, na.rm = TRUE)) |>
    group_by(recordid) |>
      mutate(
        prevarrivaldate   = lag(arrivaldate),
        prevarrivaltype   = lag(arrivaltype),
        prevarrivalreason = lag(arrivalreason),
        prevarrivalmorph  = lag(arrivalmorph),
        prevarrivalflow   = lead(arrivalflow),
        
        nextarrivaldate   = lead(arrivaldate),
        nextarrivaltype   = lead(arrivaltype),
        nextarrivalreason = lead(arrivalreason),
        nextarrivalmorph  = lead(arrivalmorph),
        nextarrivalflow   = lead(arrivalflow),
        
        # Adding ranking
        arrival = row_number(),
        
        # Count all treatment plans
        allarrivalcnt = n(),

        # Count FHCC arrivals
        fhccarrivalcnt = sum(grepl("^arr", arrivaltable))
        
        
      ) |>
    ungroup() |>
    select(recordid, ptmrn, arrival, everything())

cat('Rank Patient Arrivals with prev/next')
```

## Report Missing Arrival Sequence number or Arrival Date

Very hard to get arrivals in order without this information, luckily not very many are missing these. When patients are entered into REDCap they are assigned a sequence number on the top of each arrival form. This is useful in the interface if we need to change the order in which arrivals occur. This becomes especially handy if a treatment attempt was missed during abstraction and needs to be inserted.\
\
The consequence is that sometimes we forget to adjust all of the sequence numbers associated with the various arrivals for the patient. When this is the case this code checks for any discrepancies and creates a report.\
\
It also checks for arrival date, missing arrival dates also make it difficult to tell in what order events took place.

```{r}
# identify errors
missing_arrival_order_info <- allarrival_allprior |> 
  # bind_rows(allnotreat) |>
  left_join(patient_list |> select(recordid, patient_list_rectime))  |> 
  arrange(recordid, arrivaldate, treatmentdate, arrivalnum, arrivaltable ) |>
  filter(!grepl('^pre', arrivaltable)) |>
  filter(is.na(arrivaldate) | arrivalnum == "") |>
  filter(recordid > 1000 & arrivaltable != 'induction') |>
  arrange( ptlastname, ptmrn, arrivalnum ) |>
  select(  recordid, ptmrn, ptlastname
         , arrivalnum, arrivaltable, orig_arrivaltable, arrivaldate, arrival_rectime, arrivalnote
         , treatmentdate, treatmentnote, patient_list_rectime )

if (nrow(missing_arrival_order_info) > 0){
  # Create excel for data cleaning
  make_sheet(missing_arrival_order_info, scrapdir, 'nodate_or_nonumber', 'missing_arrival_order_info', 'xlsx', overwrite=TRUE )
  # Add to table of things to correct
  addrow('todo_df', scrapdir, 'missing_arrival_order_info.xlsx', 'Review for arrival date missing and arrival number order')
  cat('\nReport Missing Arrival Sequence number or Arrival Date')
} else {
  cat('\nNo Missing Arrival Sequence number or Arrival Date')
}
```

## SAVE TO SQLSERVER

Copy tables to SQL Server. The table **allarrivals** shows treatment for patients, including those prior to arrival here. It also includes the patients who have not been abstracted. The table **allnotreat** contains patients who were consults, comfort care, or for other reasons did not get curative focused treatment. The table **noarrival** shows patients who we have triaged for abstraction that have been registered into CTMS but have not yet had their abstraction completed, i.e. no arrival data has been completed for them.

Nuances:

I bind all the data together and make a temp table so that I can adjust it to get the correct column structure.  Then copy that structure to the tables.

After that the sql_insert_df() command actually appends the data, it does check to see if there needs to be resizing, but that is pretty much already done via the temp table, therefore the later inserts are cheaper.

The process is also sped up because when we make the dated copies it is all done server side with SQL pass through.

```{r}
# Resize accommodate structure of all the records
cummarrival <- allarrival |> bind_rows(allnotreat) |> bind_rows(allprior) |> distinct()
tempsch = 'scratch'
sql_drop_tbl(tempsch, 'temp')
sql_insert_df(cummarrival, 'scratch.temp')
# make empty copies
sql_copy_tbl(tempsch, 'allnotreat', 'temp', data=FALSE)
sql_copy_tbl(tempsch, 'allprior',   'temp', data=FALSE)
sql_copy_tbl(tempsch, 'allarrival', 'temp', data=FALSE)
sql_drop_tbl(tempsch, 'temp')

# insert allarrivals, allnotreat, and allprior twice, normal and again with a date suffix
x <- sql_insert_df(allarrival, 'frozen.allarrival', date_suffix='%Y%m')
x <- sql_insert_df(allnotreat, 'frozen.allnotreat', date_suffix='%Y%m') 
x <- sql_insert_df(allprior,   'frozen.allprior',   date_suffix='%Y%m') 

cat('SQL Server updated with allarrival, allnotreat, and allprior tables')
```
